---
title: "Navigating the electoralContestation Package"
output: html_document
---


## Build

You can build the package from `github` by installing the package `devtools` and then using that to install from my `github` repository

```{r}
#install.packages('devtools')
devtools::install_github("crweber9874/electoralContestation", force= TRUE)
```

## Data
There are sample data in repository to avoid needing to download, recode, etc from your local machine. This needs some updating, as the data build is the data for the 2025 MPSA meeting and does not yet include the newer BYU CES data
```{r}
library(electoralContestation)
library(dplyr)
library(tidybayes)
electoral_contestation %>% head()
```
## Estimation Steps

The package is meant to help with the analytic workflow. I've organized the functions thematically:

* `modeling`: There are three main modeling functions, `brms.ordinal`, `brms.linear` and `brms.nominal`. These are used to estimate ordinal, linear and nominal models, respectively. The `IV` variable is the independent variable, and the `DV` variable is the dependent variable. The `IV` variable is specified as a formula, so that higher order interactions can easily be specified. Thus far, I've only implemented the `brms.ordinal` function, so we can modify this one to our needs and then I'll update the others.

* `prediction`: There are two main prediction functions, `posterior_means` and `posterior_pme`. The `posterior_means` function generates posterior predictions with two-way interactions, while the `posterior_pme` function generates posterior predictions with marginal effects. These functions are used to generate predictions from the models estimated in the previous step.

* `plotting`: There are two main plotting functions, `ggPoint` and `ggMargins`. The `ggPoint` function generates a point estimate plot with a rug plot of the dependent variable distribution, while the `ggMargins` function generates a marginal effects plot with a rug plot of the dependent variable distribution. These functions are used to visualize the predictions genera
