---
title: "Building a Electoral Contestation Survey Repository"
output: html_document
---

If you set your own $\texttt{Sys.setenv("OPENAI_API_KEY" = "")$ you can use copilot's code completion. Otherwise, I just connect to the Dropbox folder on my computer, and read in the files, I then recode some common variables.

## AVP 2023, AVP Wave 1
## Sample Description 

**Sampling Frame**: Voters in the 2022 Election

YouGov recruited a representative sample of n=1000 Arizona resident adults to complete the survey. The respondents were matched to a sampling frame based on gender, age, race, and education. This sampling frame is intended to serve as a politically representative sample of Arizona registered voters, constructed from the American Community Survey (ACS) public use microdata file, public voter file records, the 2020 Current Population Survey (CPS) Voting and Registration supplements, the 2020 National Election Pool (NEP) exit poll, and the 2020 CES surveys, including demographics and 2020 presidential vote.

```{r, echo = FALSE, message = FALSE, warning = FALSE}
library(readstata13)
library(haven)
library(tidyverse)
library(chattr)
library(dplyr)
devtools::load_all()
# Sys.setenv("OPENAI_API_KEY" = "")
# The key will not be shown in the repository, just your local .renv environment.
setwd("/Users/Chris/Dropbox/electoral_contestation-main/Data Repository/")

wss20 <- read.dta13("2020 WSS.dta")
cesASU <- read.dta13("2022 CES ASU.dta")
cesBYU <- read.dta13("2022 CES BYU.dta")
wss24 <- read.dta13("2024 WSS.dta")
avp23 <- read_sav("UARZ0003_OUTPUT.sav")
avp24 <- read_sav("UARZ0004_OUTPUT.sav")
cesBYU24 <- read_sav("CCES24_BYU_OUTPUT.sav")

```

Recode each data set with a minimal set of common demographic variables.

```{r}
original_columns <- names(avp23)
avp23 <- avp23 %>%
  # common variables
  mutate(black = ifelse(as.numeric(race) == 2, 1, 0),
         white = ifelse(as.numeric(race) == 1, 1, 0),
         latino = ifelse(as.numeric(race) == 3, 1, 0),
         asian = ifelse(as.numeric(race) == 4, 1, 0),
         american_indian = ifelse(as.numeric(race) == 5, 1, 0),
         age = 2022 - as.numeric(birthyr),
         married = ifelse(as.numeric(marstat) == 1, 1, 0),
         female = ifelse(as.numeric(gender) == 2, 1, 0),
         college = ifelse(as.numeric(educ) >= 5, 1, 0),
         faminc = ifelse(as.numeric(faminc_new) > 8, 1, 0),
         survey_weight = weight,
         county_fips = as.numeric(county_AZ),
         CD = as.numeric(cd),
         LD = as.numeric(LD_upper),
         zipcode = as.numeric(inputzip),
         caseID = caseid, #By whatever is in the data,
         state = "4",
         year = 2023,
         survey = "avpw1",
         prepost = "post"
         )
head(avp23)
```        

I created some functions in the package to avoid repetitive code -- particularly when recoding survey items that follow a common set of response categories.  These are in the $\texttt{/R}$ folder. Some are utilities and denoted $\texttt{util_}$. In any case, here's how I recoded the 2023 AVP data file.

```{r}
recode_rules <- list(
  list(column = "presvote20post", recode_rules = c("1" = 0, "2" = 1), new_column = "presvote_trump_2020"),
  list(column = "vote_2020", recode_rules = c("1" = 1, "2" = 0), new_column = "voted2020"),
  list(column = "MIP", recode_rules = c(
    "1" = "economy",
    "2" = "climate/environment",
    "3" = "healthcare",
    "4" = "racism",
    "5" = "terrorism",
    "6" = "abortion",
    "7" = "crime",
    "8" = "poverty",
    "9" = "national security",
    "10" = "education",
    "11" = "democracy",
    "12" = "other"
  ), new_column = "most_important_problem"),
  list(column = "contestation1", recode_rules = five_r, new_column = "attend_march"),
  list(column = "contestation2", recode_rules = five_r, new_column = "criticize_election"),
  list(column = "contestation3", recode_rules = five_r, new_column = "burn_flag"),
  list(column = "contestation4", recode_rules = five_r, new_column = "recount"),
  list(column = "contestation5", recode_rules = five_r, new_column = "court"),
  list(column = "contestation6", recode_rules = five_r, new_column = "certify"),
  list(column = "contestation7", recode_rules = five_r, new_column = "concede"),
  list(column = "contestation8", recode_rules = five_r, new_column = "state_certify"),
  list(column = "contestation9", recode_rules = five_r, new_column = "violent_protest"),
  list(column = "contestation10", recode_rules = five_r, new_column = "new_election"),
  list(column = "steal_2020", recode_rules = five_r, new_column = "stolen_2020"),
  list(column = "free_and_fair", recode_rules = five_r, new_column = "free_fair"),
  list(column = "vote_confidence", recode_rules = five_r, new_column = "voteconfidence"),
  list(column = "auth1", recode_rules = c("1" = 0, "2" = 1), new_column = "auth_1"),
  list(column = "auth2", recode_rules = c("1" = 0, "2" = 1), new_column = "auth_2"),
  list(column = "auth3", recode_rules = c("1" = 1, "2" = 0), new_column = "auth_3"),
  list(column = "auth4", recode_rules = c("1" = 0, "2" = 1), new_column = "auth_4"),
  list(column = "rr1", recode_rules = four_n, new_column = "rr_1"),
  list(column = "rr2", recode_rules = four_r, new_column = "rr_2"),
  list(column = "rr3", recode_rules = four_r, new_column = "rr_3"),
  list(column = "rr4", recode_rules = four_r, new_column = "rr_4"),
  list(column = "pid7", recode_rules = c("1" = 1, "2" = 2, "3" = 3, "4" = 4, "5" = 5, "6" = 6, "7" = 7), new_column = 
         "party_identification7"),
  list(column = "religpew", recode_rules = c(
    "1" = 1, "2" = 1, "3" = 1, "4" = 1, "5" = 0, "6" = 0, "7" = 0,
    "8" = 0, "9" = 0, "10" = 0, "11" = 0, "12" = 0
  ), new_column = "christian"),
  list(column = "presvote16post", recode_rules = c("1" = 0, "2" = 0), 
       new_column = "presvote_trump_2016"),
  list(column = "ideo5", 
       recode_rules = c("1" = 1, "2" = 2, "3" = 3, "4" = 4, "5" = 5), 
       new_column = "ideology5")
)

avp23 <- recodeList(avp23, recode_rules)
new_columns <- names(avp23)
new_columns <- setdiff(new_columns, original_columns)

avp23 <- avp23 %>% select(all_of(new_columns))
```
## The 2024 AVP Data, Wave II (*n* = 803; *n$_{recontact}$* = 403)
### Description

**Sampling Frame**: Voters in the 2022 Election, including 403 recontacts from Wave I

YouGov interviewed 865 respondents and matched them down to a sample of 803 for the final dataset. Of these 803 respondents, 403 were recontacted from a previous study conducted between December 14, 2022, and December 28, 2022, AVP Wave 1. The remaining new respondents were matched to a sampling frame based on gender, age, race, and education. This sampling frame is intended to serve as a politically representative sample of Arizona registered voters, constructed from the American Community Survey (ACS) public use microdata file, public voter file records, the 2020 Current Population Survey (CPS) Voting and Registration supplements, the 2020 National Election Pool (NEP) exit poll, and the 2020 CES surveys, including demographics and 2020 presidential vote.

The matched cases were weighted to the sampling frame using propensity scores. A logistic regression was estimated for inclusion in the frame by combining the matched cases and the frame. The propensity score function included age, gender, race/ethnicity, years of education, and region. The propensity scores were grouped into deciles of the estimated propensity score in the frame and post-stratified according to these deciles. The weights were then post-stratified based on the 2020 presidential vote choice and a four-way stratification of gender, age (4 categories), race (4 categories), and education (4 categories) to produce the final weight.

```{r}
original_columns <- names(avp24)

avp24 <- avp24 %>%
  mutate(black = ifelse(as.numeric(race) == 2, 1, 0),
         white = ifelse(as.numeric(race) == 1, 1, 0),
         latino = ifelse(as.numeric(race) == 3, 1, 0),
         asian = ifelse(as.numeric(race) == 4, 1, 0),
         american_indian = ifelse(as.numeric(race) == 5, 1, 0),
         age = 2022 - as.numeric(birthyr),
         married = ifelse(as.numeric(marstat) == 1, 1, 0),
         college = ifelse(as.numeric(educ) >= 5, 1, 0),
         faminc = ifelse(as.numeric(faminc_new) > 8, 1, 0),
         survey_weight = weight,
         county_fips = as.numeric(county_AZ),
         CD = as.numeric(CD118),
         LD = as.numeric(LD_U),
         zipcode = as.numeric(inputzip),
         caseidID = caseid,
         caseidID22 = UARZ0003_caseid,
         state = "4",
         year = 2024,
         survey = "avpw2",
         prepost = "post"

         )
## Code Survey items
recode_rules <- list(
  list(column = "voteChoice", recode_rules = c("1" = 0, "2" = 1), new_column = "presvote_trump_2020"),
  list(column = "MIP", recode_rules = c(
    "1" = "economy",
    "2" = "climate/environment",
    "3" = "climate/enviornment",
    "5" =  "racism",
    "6" = "terrorism",
    "7" = "abortion",
    "8" = "crime",
    "9" = "poverty",
    "10" = "national security",
    "11" = "education",
    "12" = "democracy",
    "13" = "immigration",
    "14" = "guns",
    "15" = "foreign_policy",
    "16" = "other"),  new_column = "most_important_problem"),
  list(column = "contestation1", recode_rules = five_r,   new_column = "attend_march"),
  list(column = "contestation2", recode_rules = five_r,   new_column = "criticize_election"),
  list(column = "contestation3", recode_rules = five_r,   new_column = "burn_flag"),
  list(column = "contestation4", recode_rules = five_r,   new_column = "recount"),
  list(column = "contestation5", recode_rules = five_r,   new_column = "court"),
  list(column = "contestation6", recode_rules = five_r,   new_column = "certify"),
  list(column = "contestation7", recode_rules = five_r,   new_column = "concede"),
  list(column = "contestation8", recode_rules = five_r,   new_column = "state_certify"),
  list(column = "contestation9", recode_rules = five_r,   new_column = "violent_protest"),
  list(column = "contestation10", recode_rules = five_r,  new_column = "new_election"),
  list(column = "auth1", recode_rules = c("1" = 0, "2" = 1),   new_column = "auth_1"),
  list(column = "auth2", recode_rules = c("1" = 0, "2" = 1),   new_column = "auth_2"),
  list(column = "auth3", recode_rules = c("1" = 1, "2" = 0),   new_column = "auth_3"),
  list(column = "auth4", recode_rules = c("1" = 0, "2" = 1),   new_column = "auth_4"),
  list(column = "pid7", recode_rules = c("1" = 1, "2" = 1, "3" = 1, "4" = 2,
                                           "5" = 3, "6" = 3, "7" = 3),   new_column = "party_identification3"),
  list(column = "pid7", recode_rules = c("1" = 1, "2" = 2, "3" = 3, "4" = 4, "5" = 5, "6" = 6, "7" = 7), new_column = 
         "party_identification7"),
  list(column = "religpew", recode_rules = c("1" = 1, "2" = 1, "3" = 1, "4" = 1,
                                               "5"= 0, "6" = 0, "7" = 0, "8" = 0, "9" = 0, 
                                               "10" = 0, "11" = 0, "12" = 0),   new_column =
         "christian"),
  list(column = "ideo5", recode_rules = c("1" = 1, "2" = 2, "3" = 3, "4" = 4, "5" = 5),  
       new_column = "ideology5"),
  list(column = "gender4", recode_rules = c("2" = 1, "1" = 0),   new_column = "female"))
# Function to iterates  
# Replace: devtools::load_all()
avp24 = recodeList(avp24, recode_rules) 
new_columns <- names(avp24)
new_columns <- setdiff(new_columns, original_columns)

# Extract the recoded columns
avp24 <- avp24 %>% select(all_of(new_columns))
```   

## Western States
## 2020

**Sampling Frame**: Voters across five western states, Latino oversample

YouGov interviewed 3577 respondents who were then matched down to a
sample of 3000 to produce the final dataset. An oversample of 600
Latinos were interviewed for a total of 3600. The respondents were
matched to a sampling frame on gender, age, race, and education. The
frame was constructed by stratified sampling from the full 2018 American
Community Survey (ACS) 1-year sample with selection within strata by weighted
sampling with replacements (using the person weights on the public use
file).

The matched cases were weighted to the sampling frame using propensity
scores. The matched cases and the frame were combined and a logistic
regression was estimated for inclusion in the frame. The propensity
score function included age, gender, race/ethnicity, years of education,
and region. The propensity scores were grouped into deciles of the
estimated propensity score in the frame and post-stratified according to
these deciles.

The weights were then post-stratified on 2016 Presidential vote choice, and
a four-way stratification of gender, age (4-categories), race (4-
categories), and education (4-categories), to produce the final weight.

The empirical design is a natural experiment; comparable numbers of participants 
complete pre or post election, or interregnum period. Five "core" items are present

$\bullet$ Attend march or demonstration
$\bullet$ Burn the American flag
$\bullet$ Contest the outcome in the courts
$\bullet$ Support ballot recounts
$\bullet$ Publicly criticize the integrity or fairness of the election

Two questions are experimentally varied.

$\bullet$ Attend march or demonstration [“even if turns violent”]
$\bullet$ Publicly criticize the integrity or fairness of the election ["on social media"]

Key findings: This is an interrupted design showing treatment effects within days of the election, large number of psychological measures, large number of trust/participation variables that can be used to distinguish contestation from related constructs



```{r}
## Code common items
original_columns <- names(wss20)

wss20 <- wss20 %>%
  mutate(black = ifelse(as.numeric(race) == 2, 1, 0),
         white = ifelse(as.numeric(race) == 1, 1, 0),
         latino = ifelse(as.numeric(race) == 3, 1, 0),
         asian = ifelse(as.numeric(race) == 4, 1, 0),
         american_indian = ifelse(as.numeric(race) == 5, 1, 0),
         age = 2022 - birthyr,
         married = ifelse(as.numeric(marstat) == 1, 1, 0),
         college = ifelse(as.numeric(educ) >= 5, 1, 0),
         faminc = ifelse(as.numeric(faminc_new) > 8, 1, 0),
         survey_weight = weight,
         caseidID = caseid,
         state = as.character(inputstate),
         year = 2020,
         survey = "wss20",
         DATE = as.Date(endtime, "%YY-%mm-%dd", tz = "MST"),
         post_election = ifelse(DATE > "2020-11-03", 1, 0),
         post_call = ifelse(DATE >= "2020-11-07", 1, 0),
         uncertainty = ifelse(DATE > "2020-11-03", 1, 0),
         prepost = ifelse(DATE > "2020-11-03", "pre", "post")

  )

       
## Code Survey items
recode_rules <- list(
  list(column = "WSS15s", recode_rules = c(
    "1" = "environment",
    "2" = "police",
    "3" = "economy",
    "5" =  "housing",
    "6" = "trade/foreign policy",
    "7" = "trade/foreign policy",
    "8" = "education",
    "9" = "healthcare",
    "10" = "covid",
    "11" = "other"),  new_column = "most_important_problem"),
  list(column = "WSS32_1", recode_rules = c("1" = 1, "2" = 0),   new_column = "polMeeting"),
  list(column = "WSS32_2", recode_rules = c("1" = 1, "2" = 0),   new_column = "polSign"),
  list(column = "WSS32_3", recode_rules = c("1" = 1, "2" = 0),   new_column = "polVolunteer"),
  list(column = "WSS32_4", recode_rules = c("1" = 1, "2" = 0),   new_column = "polProtest"),
  list(column = "WSS32_5", recode_rules = c("1" = 1, "2" = 0),   new_column = "polOfficial"),
  list(column = "WSS32_6", recode_rules = c("1" = 1, "2" = 0),   new_column = "polDonate"),
  list(column = "WSS32_7", recode_rules = c("1" = 1, "2" = 0),   new_column = "polSocial"),
  list(column = "WSS32_8", recode_rules = c("1" = 1, "2" = 0),   new_column = "polPersuade"),
  list(column = "WSS32_9", recode_rules = c("1" = 1, "2" = 0),   new_column = "polNone"),
  list(column = "WSS54_1", recode_rules = four_r,   new_column = "trustCongress"),
  list(column = "WSS54_2", recode_rules = four_r,   new_column = "trustPresident"),
  list(column = "WSS54_3", recode_rules = four_r,   new_column = "trustSC"),
  list(column = "WSS54_4", recode_rules = four_r,   new_column = "trustGovernment"),
  list(column = "WSS54_5", recode_rules = four_r,   new_column = "trustStateleg"),
  list(column = "WSS54_6", recode_rules = four_r,   new_column = "trustPolice"),
  list(column = "WSS54_7", recode_rules = four_r,   new_column = "trustScience"),
  list(column = "WSS33_1", recode_rules = five_n,   new_column = "internal_efficacy"),
  list(column = "WSS33_2", recode_rules = five_n,   new_column = "external_efficacy"),
  list(column = "WSS40_1_split", recode_rules = c("1" = 1, "2" = 0),   new_column = "violent_treat"),
  list(column = "WSS40_5_split", recode_rules = c("1" = 1, "2" = 0),   new_column = "socialmedia_treat"),
  list(column = "WSS36_b", recode_rules = c("1" = 1, "2" = 0),   new_column = "v1"),
  list(column = "WSS36_c", recode_rules = c("1" = 1, "2" = 0),   new_column = "v2"),
  list(column = "WSS40_1", recode_rules = five_r,   new_column = "attend_march"),
  list(column = "WSS40_2", recode_rules = five_r,   new_column = "criticize_election"),
  list(column = "WSS40_3", recode_rules = five_r,   new_column = "burn_flag"),
  list(column = "WSS40_4", recode_rules = five_r,   new_column = "recount"),
  list(column = "WSS40_5", recode_rules = five_r,   new_column = "court"),
  list(column = "WSS07_1a", recode_rules = c("1" = 0, "2" = 1),   new_column = "auth_1"),
  list(column = "WSS07_1b", recode_rules = c("1" = 1, "2" = 0),   new_column = "auth_2"),
  list(column = "WSS07_1c", recode_rules = c("1" = 0, "2" = 1),   new_column = "auth_3"),
  list(column = "WSS07_1d", recode_rules = c("1" = 0, "2" = 1),   new_column = "auth_4"),
  list(column = "pid7", recode_rules = c("1" = 1, "2" = 1, "3" = 1, "4" = 2,
                                           "5" = 3, "6" = 3, "7" = 3),   new_column = "party_identification3"),
  list(column = "pid7", recode_rules = c("1" = 1, "2" = 2, "3" = 3, "4" = 4, "5" = 5, "6" = 6, "7" = 7), new_column = 
         "party_identification7"),
  list(column = "religpew", recode_rules = c("1" = 1, "2" = 1, "3" = 1, "4" = 1,
                                               "5"= 0, "6" = 0, "7" = 0, "8" = 0, "9" = 0, 
                                               "10" = 0, "11" = 0, "12" = 0),   new_column =
                                             "christian"),
  list(column = "ideo5", recode_rules = c("1" = 1, "2" = 2, "3" = 3, "4" = 4, "5" = 5),  
       new_column = "ideology5"),
  list(column = "gender", recode_rules = c("2" = 1, "1" = 0),   new_column = "female"))
# Function to iterates  
# Replace: devtools::load_all()
wss20 = recodeList(wss20, recode_rules) 

wss20$violent_march <- ifelse(wss20$WSS40_1_split == 1, wss20$attend_march, NA)
wss20$attend_march <-  ifelse(wss20$WSS40_1_split == 0, wss20$attend_march, NA)

wss20$criticize_election <- ifelse(wss20$WSS40_5_split == 0, wss20$criticize_election, NA)
wss20$criticize_election_social_media <- ifelse(wss20$WSS40_1_split == 1, wss20$criticize_election, NA)


wss20$presvote_trump_2020 <- ifelse(wss20$WSS36_b == 9, wss20$v2, wss20$v1)
new_columns <- names(wss20)
new_columns <- setdiff(new_columns, original_columns)

# Extract the recoded columns
wss20 <- wss20 %>% select(all_of(new_columns))
```  


## Western States
## 2024

**Sampling Frame**: Voters across five western states, Latino oversample

YouGov interviewed 5082 respondents pre and post the 2024 election across the
states of Arizona (1286), Colorado (1187), Nevada (615), New Mexico (664), and
Utah (618), in addition to an oversample of 715 Hispanic respondents. 

The Hispanic oversample was matched down to 600 observations (300 before and after
the election), matched to a sampling frame on gender, age, race, and education.
The sampling frame is a politically representative "modeled frame" of Hispanic
US adults, based upon the American Community Survey (ACS) public use microdata
file, public voter file records, the 2020 Current Population Survey (CPS)
Voting and Registration supplements, the 2020 National Election Pool (NEP) exit
poll, and the 2020 CES surveys, including demographics and 2020 presidential vote.
The matched cases were weighted to the sampling frame using propensity scores.
The matched cases and the frame were combined and a logistic regression was
estimated for inclusion in the frame. The propensity score function included age,
gender, race/ethnicity, years of education, and state. The propensity scores were
grouped into deciles of the estimated propensity score in the frame and
post-stratified according to these deciles.

The weights were then post-stratified on 2020 presidential vote choice as well
as a four-way stratification of gender, age (4-categories), race (4-categories),
and education (4-categories), to produce the Hispanic over sample weight.
We next proceeded to match down samples by state -- initially excluding the
Hispanic oversample results -- with 1100 for Arizona (550 before/after), 1000
for Colorado (500 before/after), 450 for Nevada (225 before/after), 400 for
New Mexico (200 before/after), and 450 for Utah (225 before/after). The respondents
were matched to a sampling frame on gender, age, race, and education by state.
The sampling frame is a politically representative "modeled frame" of US adults
within these states, based upon the American Community Survey (ACS) public use
microdata file, public voter file records, the 2020 Current Population Survey
(CPS) Voting and Registration supplements, the 2020 National Election Pool (NEP)
exit poll, and the 2020 CES surveys, including demographics and 2020 presidential
vote.

The matched cases were weighted to the sampling frame using propensity scores.
The matched cases and the frame were combined and a logistic regression was
estimated for inclusion in the frame. The propensity score function included age,
gender, race/ethnicity, years of education, and home ownership. The propensity
scores were grouped into deciles of the estimated propensity score in the frame
and post-stratified according to these deciles.
The weights were then post-stratified on 2020 presidential vote choice, 2024
presidential vote choice, as well as a four-way stratification of gender, age
(4-categories), race (4-categories), and education (4-categories), to produce
the initial weights by state.

Upon acquiring the initial weights by state, the respective Latino oversample
observations were added into their respective states. These aggregated state
frames were then post stratified by the Hispanic proportion of the population
within these states, as determined via the sampling frame to produce the final
weights by state.

Finally, all of the state observations were aggregated, with results weighted by
the proportion of the state's population within the sampling frame, to produce
a final single aggregated weight for the final 4000 observations.

```{r}
original_columns <- names(wss24)

wss24 <- wss24 %>%
  mutate(black = ifelse(as.numeric(race) == 2, 1, 0),
         white = ifelse(as.numeric(race) == 1, 1, 0),
         latino = ifelse(as.numeric(race) == 3, 1, 0),
         asian = ifelse(as.numeric(race) == 4, 1, 0),
         american_indian = ifelse(as.numeric(race) == 5, 1, 0),
         female = ifelse(gender == 2, 1, 0),
         age = 2024 - birthyr,
         married = ifelse(as.numeric(marstat) == 1, 1, 0),
         college = ifelse(as.numeric(educ) >= 5, 1, 0),
         faminc = ifelse(as.numeric(faminc_new) > 8, 1, 0),
         survey_weight = weight_all,
         caseidID = caseid,
         state = as.character(inputstate),
         WSS007_b = as.numeric(WSS007_b),
         WSS007_c = as.numeric(WSS007_c),
         year = 2024,
         survey = "wss24",
         pre_post_election = ifelse(wave == 1, 1, 0),
         # 0 - 100
         trustPresident = as.numeric(WSS022_a)/100,
         trustCongress = as.numeric(WSS022_b)/100,
         trustSC = as.numeric(WSS022_c)/100,
         trustGovernment = as.numeric(WSS022_d)/100,
         trustStateleg = as.numeric(WSS022_e)/100,
         )


## Code Survey items
recode_rules <- list(
  list(column = "WSS009_a", recode_rules = five_n,   new_column = "internal_efficacy"),
  list(column = "WSS009_b", recode_rules = five_n,   new_column = "external_efficacy"),
  list(column = "WSS007_b", recode_rules = c("1" = 1, "2" = 0),   new_column = "v1"),
  list(column = "WSS007_c", recode_rules = c("1" = 1, "2" = 0),   new_column = "v2"),
  list(column = "WSS029_a", recode_rules = five_r,   new_column = "attend_march"),
  list(column = "WSS029_b", recode_rules = five_r,   new_column = "burn_flag"),
  list(column = "WSS029_c", recode_rules = five_r,   new_column = "court"),
  list(column = "WSS029_d", recode_rules = five_r,   new_column = "recount"),
  list(column = "WSS029_e", recode_rules = five_r,   new_column = "criticize_election"),
  list(column = "WSS029_f", recode_rules = five_r,   new_column = "concede"),
  list(column = "WSS029_g", recode_rules = five_r,   new_column = "occupy_buildings"),
  list(column = "WSS029_h", recode_rules = five_r,   new_column = "object_eligibility"),
  list(column = "WSS013_a", recode_rules = c("1" = 0, "2" = 1),   new_column = "auth_1"),
  list(column = "WSS013_b", recode_rules = c("1" = 1, "2" = 0),   new_column = "auth_2"),
  list(column = "WSS013_c", recode_rules = c("1" = 0, "2" = 1),   new_column = "auth_3"),
  list(column = "WSS013_d", recode_rules = c("1" = 0, "2" = 1),   new_column = "auth_4"),
  list(column = "pid7", recode_rules = c("1" = 1, "2" = 1, "3" = 1, "4" = 2,
                                           "5" = 3, "6" = 3, "7" = 3),   new_column = "party_identification3"),
  list(column = "pid7", recode_rules = c("1" = 1, "2" = 2, "3" = 3, "4" = 4, "5" = 5, "6" = 6, "7" = 7), new_column = 
         "party_identification7"),  
  
  list(column = "religpew", recode_rules = c("1" = 1, "2" = 1, "3" = 1, "4" = 1,
                                               "5"= 0, "6" = 0, "7" = 0, "8" = 0, "9" = 0, 
                                               "10" = 0, "11" = 0, "12" = 0),   new_column =
                                             "christian"),
  list(column = "ideo5", recode_rules = c("1" = 1, "2" = 2, "3" = 3, "4" = 4, "5" = 5),  
       new_column = "ideology5"),
  list(column = "presvote20post", recode_rules = c("1" = 0, "2" = 1),  
       new_column = "presvote_trump_2020"),
  list(column = "presvote16post", recode_rules = c("1" = 0, "2" = 1),  
       new_column = "presvote_trump_2016")
)
# Function to iterates  
# Replace: devtools::load_all()
wss24 = recodeList(wss24, recode_rules) 
wss24$presvote_trump_2024 <- ifelse(is.na(wss24$v1), wss24$v2, wss24$v1)
new_columns <- names(wss24)
new_columns <- setdiff(new_columns, original_columns)

# Extract the recoded columns
wss24 <- wss24 %>% select(all_of(new_columns))
```  

## CES ASU Module

Pretest, followed with post-test, randomization to conditions (50, 25, 25).

Pretest: "Many people are unhappy with the outcomes of elections. How much do you support or oppose each of the following behaviors when people are unhappy with the outcome of an election?"

Control: "Many people are unhappy with the outcomes of elections. How much do you support or oppose each of the following behaviors when people are unhappy with the outcome of an election?"


Fraud: Many people are unhappy with the outcomes of elections, especially when there are concerns about election fraud. How much do you support or oppose each of the following behaviors when people are unhappy with the outcome of an election? 

Norms: The peaceful transfer of power is a foundation of American democracy. However, many people are unhappy with the outcomes of elections. How much do you support or oppose each of the following behaviors when people are unhappy with the outcome of an election? 

```{r}
original_columns <- names(cesASU)

cesASU <- cesASU %>%
  mutate(black = ifelse(as.numeric(race) == 2, 1, 0),
         white = ifelse(as.numeric(race) == 1, 1, 0),
         latino = ifelse(as.numeric(race) == 3, 1, 0),
         asian = ifelse(as.numeric(race) == 4, 1, 0),
         american_indian = ifelse(as.numeric(race) == 5, 1, 0),
         age = 2023 - birthyr,
         married = ifelse(as.numeric(marstat) == 1, 1, 0),
         college = ifelse(as.numeric(educ) >= 5, 1, 0),
         faminc = ifelse(as.numeric(faminc_new) > 8 & as.numeric(faminc_new) < 100, 1, 0),
         caseidID = caseid,
         state = as.character(inputstate),
         year = 2022,
         survey = "cesASU"
)
    
## Code Survey items -- follows pre-post test design
recode_rules <- list(
  list(column = "CC22_430a_1", recode_rules = c("1" = 1, "2" = 0),   new_column = "polMeeting"),
  list(column = "CC22_430a_2", recode_rules = c("1" = 1, "2" = 0),   new_column = "polSign"),
  list(column = "CC22_430a_3", recode_rules = c("1" = 1, "2" = 0),   new_column = "polVolunteer"),
  list(column = "CC22_430a_4", recode_rules = c("1" = 1, "2" = 0),   new_column = "polProtest"),
  list(column = "CC22_430a_5", recode_rules = c("1" = 1, "2" = 0),   new_column = "polOfficial"),
  list(column = "CC22_430a_6", recode_rules = c("1" = 1, "2" = 0),   new_column = "polDonate"),
  list(column = "CC22_430a_8", recode_rules = c("1" = 1, "2" = 0),   new_column = "polNone"),
  # Pre Election
  list(column = "ASU351", recode_rules = five_n,   new_column = "attend_march"),
  list(column = "ASU352", recode_rules = five_n,   new_column = "burn_flag"),
  list(column = "ASU353", recode_rules = five_n,   new_column = "court"),
  list(column = "ASU354", recode_rules = five_n,   new_column = "recount"),
  list(column = "ASU355", recode_rules = five_n,   new_column = "criticize_election"),
  list(column = "ASU356", recode_rules = five_n,   new_column = "violent_acts"),
  # Control
  list(column = "ASU412", recode_rules = five_r,   new_column = "attend_march1"),
  list(column = "ASU413", recode_rules = five_r,   new_column = "burn_flag1"),
  list(column = "ASU414", recode_rules = five_r,   new_column = "court1"),
  list(column = "ASU415", recode_rules = five_r,   new_column = "recount1"),
  list(column = "ASU416", recode_rules = five_r,   new_column = "criticize_election1"),
  list(column = "ASU417", recode_rules = five_r,   new_column = "violent_acts1"),
  # Fraud
  list(column = "ASU418", recode_rules = five_r,   new_column = "attend_march2"),
  list(column = "ASU419", recode_rules = five_r,   new_column = "burn_flag2"),
  list(column = "ASU420", recode_rules = five_r,   new_column = "court2"),
  list(column = "ASU421", recode_rules = five_r,   new_column = "recount2"),
  list(column = "ASU422", recode_rules = five_r,   new_column = "criticize_election2"),
  list(column = "ASU423", recode_rules = five_r,   new_column = "violent_acts2"),
  # Norms
  list(column = "ASU424", recode_rules = five_r,   new_column = "attend_march3"),
  list(column = "ASU425", recode_rules = five_r,   new_column = "burn_flag3"),
  list(column = "ASU426", recode_rules = five_r,   new_column = "court3"),
  list(column = "ASU427", recode_rules = five_r,   new_column = "recount3"),
  list(column = "ASU428", recode_rules = five_r,   new_column = "criticize_election3"),
  list(column = "ASU429", recode_rules = five_r,   new_column = "violent_acts3"),

  list(column = "gender4", recode_rules = c("1" = 0, "2" = 1),   new_column = "female"),
  list(column = "pid7", recode_rules = c("1" = 1, "2" = 1, "3" = 1, "4" = 2,
                                           "5" = 3, "6" = 3, "7" = 3),   new_column = "party_identification3"),
    list(column = "pid7", recode_rules = c("1" = 1, "2" = 2, "3" = 3, "4" = 4, "5" = 5, "6" = 6, "7" = 7), new_column = 
         "party_identification7"),
  list(column = "religpew", recode_rules = c("1" = 1, "2" = 1, "3" = 1, "4" = 1,
                                               "5"= 0, "6" = 0, "7" = 0, "8" = 0, "9" = 0, 
                                               "10" = 0, "11" = 0, "12" = 0),   new_column =
                                             "christian"),
  list(column = "ideo5", recode_rules = c("1" = 1, "2" = 2, "3" = 3, "4" = 4, "5" = 5),  
       new_column = "ideology5"),
  list(column = "presvote20post", recode_rules = c("1" = 0, "2" = 1),  
       new_column = "presvote_trump_2020"),
  list(column = "presvote16post", recode_rules = c("1" = 0, "2" = 1),  
       new_column = "presvote_trump_2016")
)

cesASU = recodeList(cesASU, recode_rules) 


cesASU$post_experimental_condition <- ifelse(cesASU$QASU412rand == 1, "Control", 
                                      ifelse(cesASU$QASU412rand == 2, "Fraud", 
                                          ifelse(cesASU$QASU412rand == 3, "Norms",  
                                              NA)))
  
  
  
## Kind of strange coding of the conditions, combine to common variable
cesASU = cesASU %>%
  mutate(
    attend_march_post = ifelse(!is.na(attend_march1), attend_march1,
       ifelse(!is.na(attend_march2), attend_march2,
       ifelse(!is.na(attend_march3), attend_march3,
              NA))),
    burn_flag_post = ifelse(!is.na(burn_flag1), burn_flag1,
       ifelse(!is.na(burn_flag2), burn_flag2,
       ifelse(!is.na(burn_flag3), burn_flag3,
              NA))),
    court_post = ifelse(!is.na(court1), court1,
       ifelse(!is.na(court2), court2,
       ifelse(!is.na(court3), court3,
              NA))),
    recount_post = ifelse(!is.na(recount1), recount1,
       ifelse(!is.na(recount2), recount2,
       ifelse(!is.na(recount3), recount3,
              NA))),
    criticize_election_post = ifelse(!is.na(criticize_election1), criticize_election1,
       ifelse(!is.na(criticize_election2), criticize_election2,
       ifelse(!is.na(criticize_election3), criticize_election3,
              NA)))
    )
new_columns <- names(cesASU)
new_columns <- setdiff(new_columns, original_columns)
cesASU <- cesASU %>% select(all_of(new_columns))
#
cesASU %>%
    pivot_longer(
        cols = starts_with(c("attend_march", "criticize_election")),
        names_to = "week",
        names_prefix = "EC",
        values_to = "rank",
        values_drop_na = TRUE
    ) %>% View()
```  


## CES BYU Module

```{r}
# Save original column names
original_columns <- names(cesBYU)

cesBYU <- cesBYU %>%
  mutate(black = ifelse(as.numeric(race) == 2, 1, 0),
         white = ifelse(as.numeric(race) == 1, 1, 0),
         latino = ifelse(as.numeric(race) == 3, 1, 0),
         asian = ifelse(as.numeric(race) == 4, 1, 0),
         american_indian = ifelse(as.numeric(race) == 5, 1, 0),
         age = 2023 - birthyr,
         married = ifelse(as.numeric(marstat) == 1, 1, 0),
         college = ifelse(as.numeric(educ) >= 5, 1, 0),
         faminc = ifelse(as.numeric(faminc_new) > 8 & as.numeric(faminc_new) < 100, 1, 0),
         caseidID = caseid,
         state = as.character(inputstate),
         experimental_condition_post = ifelse(BYU411_rand == "BYU4Xa", "Control", 
                                      ifelse(BYU411_rand == "BYU4Xb", "Fraud", 
                                          ifelse(BYU411_rand == "BYU4Xc", "Norms",  
                                                 NA))),
         experimental_condition_pre = ifelse(BYU311_rand == "Option A: Support or Oppose", "support_condition",
                                             ifelse(BYU311_rand == "Option B: Participate", "participate_condition",
                                                    NA)),
         year = 2022,
         survey = "cesBYU"

)
## Code Survey items
recode_rules <- list(
  list(column = "BYU767_1", recode_rules = c("1" = 1, "2" = 0),   new_column = "polButton"),
  list(column = "BYU767_2", recode_rules = c("1" = 1, "2" = 0),   new_column = "polSign"),
  list(column = "BYU767_3", recode_rules = c("1" = 1, "2" = 0),   new_column = "polPersuade"),
  list(column = "BYU767_4", recode_rules = c("1" = 1, "2" = 0),   new_column = "polTalk"),
  list(column = "BYU767_5", recode_rules = c("1" = 1, "2" = 0),   new_column = "polEmail"),
  list(column = "BYU767_6", recode_rules = c("1" = 1, "2" = 0),   new_column = "polMeme"),
  list(column = "BYU767_7", recode_rules = c("1" = 1, "2" = 0),   new_column = "polMeme_create"),
  list(column = "BYU90", recode_rules = c("1" = 1, "2" = 1, "3" = 2,
                                             "4" = 3, "5" = 4, "6" = 5,
                                             "7" = 5),   new_column = "internal_efficacy"),
  list(column = "BYU100N", recode_rules = c("1" = 1, "2" = 1, "3" = 2,
                                             "4" = 3, "5" = 4, "6" = 5,
                                             "7" = 5),   new_column = "external_efficacy"),

  # The pretest, conditions 1 and 2
  list(column = "BYU311a", recode_rules = five_n,   new_column = "attend_march1"),
  list(column = "BYU312a", recode_rules = five_n,   new_column = "burn_flag1"),
  list(column = "BYU313a", recode_rules = five_n,   new_column = "court1"),
  list(column = "BYU314a", recode_rules = five_n,   new_column = "recount1"),
  list(column = "BYU315a", recode_rules = five_n,   new_column = "criticize_election1"),
  list(column = "BYU316a", recode_rules = five_n,   new_column = "violent_acts1"),
  list(column = "BYU311b", recode_rules = five_n,   new_column = "attend_march2"),
  list(column = "BYU312b", recode_rules = five_n,   new_column = "burn_flag2"),
  list(column = "BYU313b", recode_rules = five_n,   new_column = "court2"),
  list(column = "BYU314b", recode_rules = five_n,   new_column = "recount2"),
  list(column = "BYU315b", recode_rules = five_n,   new_column = "criticize_election2"),
  list(column = "BYU316b", recode_rules = five_n,   new_column = "violent_acts2"),
  # And there is an indicator called BYU311a_rand
# This constitutes the post test  
  list(column = "BYU411", recode_rules = five_r,   new_column = "attend_marchp"),
  list(column = "BYU412", recode_rules = five_r,   new_column = "burn_flagp"),
  list(column = "BYU413", recode_rules = five_r,   new_column = "courtp"),
  list(column = "BYU414", recode_rules = five_r,   new_column = "recountp"),

# And there is another variable called BYU411_rand

  list(column = "gender4", recode_rules = c("1" = 0, "2" = 1),   new_column = "female"),
  list(column = "pid7", recode_rules = c("1" = 1, "2" = 1, "3" = 1, "4" = 2,
                                           "5" = 3, "6" = 3, "7" = 3),   new_column = "party_identification3"),
  list(column = "pid7", recode_rules = c("1" = 1, "2" = 2, "3" = 3, "4" = 4, "5" = 5, "6" = 6, "7" = 7), new_column = 
         "party_identification7"),
  list(column = "religpew", recode_rules = c("1" = 1, "2" = 1, "3" = 1, "4" = 1,
                                               "5"= 0, "6" = 0, "7" = 0, "8" = 0, "9" = 0, 
                                               "10" = 0, "11" = 0, "12" = 0),   new_column =
                                             "christian"),
  list(column = "ideo5", recode_rules = c("1" = 1, "2" = 2, "3" = 3, "4" = 4, "5" = 5),  
       new_column = "ideology5"),
  list(column = "presvote20post", recode_rules = c("1" = 0, "2" = 1),  
       new_column = "presvote_trump_2020"),
  list(column = "presvote16post", recode_rules = c("1" = 0, "2" = 1),  
       new_column = "presvote_trump_2016")
)
cesBYU = recodeList(cesBYU, recode_rules) 
# Only save the newly created ones


cesBYU %>%
  mutate(attend_march = ifelse( experimental_condition_post == "Control", attend_march1, NA),
         burn_flag = ifelse( experimental_condition_post == "Control", attend_march1, NA),
         recount = ifelse( experimental_condition_post == "Control", attend_march1, NA),
         criticize_election = ifelse( experimental_condition_post == "Control", attend_march1, NA),
         attend_march = ifelse( experimental_condition_post == "Control", attend_march1, NA),
         court = ifelse( experimental_condition_post == "Control", attend_march1, NA)) -> cesBYU
         
new_columns <- names(cesBYU)
new_columns <- setdiff(new_columns, original_columns)
cesBYU <- cesBYU %>% 
  select(all_of(new_columns))
#
```

```{r}
original_columns <- names(cesBYU24)

cesBYU <- cesBYU %>%
  mutate(black = ifelse(as.numeric(race) == 2, 1, 0),
         white = ifelse(as.numeric(race) == 1, 1, 0),
         latino = ifelse(as.numeric(race) == 3, 1, 0),
         asian = ifelse(as.numeric(race) == 4, 1, 0),
         american_indian = ifelse(as.numeric(race) == 5, 1, 0),
         age = 2025 - birthyr,
         married = ifelse(as.numeric(marstat) == 1, 1, 0),
         college = ifelse(as.numeric(educ) >= 5, 1, 0),
         faminc = ifelse(as.numeric(faminc_new) > 8 & as.numeric(faminc_new) < 100, 1, 0),
         caseidID = caseid,
         state = as.character(inputstate),
         year = 2025,
         survey = "cesBYU25"

)
## Code Survey items
recode_rules <- list(
  list(column = "CC24_430a_1", recode_rules = c("1" = 1, "2" = 0),   new_column = "polMeeting"),
  list(column = "CC24_430a_2", recode_rules = c("1" = 1, "2" = 0),   new_column = "polSign"),
  list(column = "CC24_430a_3", recode_rules = c("1" = 1, "2" = 0),   new_column = "polVolunteer"),
  list(column = "CC24_430a_4", recode_rules = c("1" = 1, "2" = 0),   new_column = "polProtest"),
  list(column = "CC24_430a_5", recode_rules = c("1" = 1, "2" = 0),   new_column = "polOfficial"),
  list(column = "CC24_430a_6", recode_rules = c("1" = 1, "2" = 0),   new_column = "polDonate"),
  list(column = "CC24_430a_8", recode_rules = c("1" = 1, "2" = 0),   new_column = "polNone"),

  # Pre and Post questions

  # The pretest, conditions 1 and 2
  list(column = "BYU400", recode_rules = five_n,   new_column = "attend_march_post"),
  list(column = "BYU401", recode_rules = five_n,   new_column = "burn_flag_post"),
  list(column = "BYU402", recode_rules = five_n,   new_column = "court_post"),
  list(column = "BYU403", recode_rules = five_n,   new_column = "recount_post"),
  list(column = "BYU404", recode_rules = five_n,   new_column = "criticize_election_post"),
  list(column = "BYU405", recode_rules = five_n,   new_column = "concede_election_post"),
  list(column = "BYU406", recode_rules = five_n,   new_column = "occupy_buildings_post"),
  list(column = "BYU407", recode_rules = five_n,   new_column = "object_eligibility_post"),
  list(column = "BYU300", recode_rules = five_n,   new_column = "attend_march_post"),
  list(column = "BYU301", recode_rules = five_n,   new_column = "burn_flag_post"),
  list(column = "BYU302", recode_rules = five_n,   new_column = "court_post"),
  list(column = "BYU303", recode_rules = five_n,   new_column = "recount_post"),
  list(column = "BYU304", recode_rules = five_n,   new_column = "criticize_election_post"),
  list(column = "BYU305", recode_rules = five_n,   new_column = "concede_election_post"),
  list(column = "BYU306", recode_rules = five_n,   new_column = "occupy_buildings_post"),
  list(column = "BYU307", recode_rules = five_n,   new_column = "object_eligibility_post"),

  # Start here
  list(column = "gender4", recode_rules = c("1" = 0, "2" = 1),   new_column = "female"),
  list(column = "pid7", recode_rules = c("1" = 1, "2" = 1, "3" = 1, "4" = 2,
                                           "5" = 3, "6" = 3, "7" = 3),   new_column = "party_identification3"),
  list(column = "pid7", recode_rules = c("1" = 1, "2" = 2, "3" = 3, "4" = 4, "5" = 5, "6" = 6, "7" = 7), new_column = 
         "party_identification7"),
  list(column = "religpew", recode_rules = c("1" = 1, "2" = 1, "3" = 1, "4" = 1,
                                               "5"= 0, "6" = 0, "7" = 0, "8" = 0, "9" = 0, 
                                               "10" = 0, "11" = 0, "12" = 0),   new_column =
                                             "christian"),
  list(column = "ideo5", recode_rules = c("1" = 1, "2" = 2, "3" = 3, "4" = 4, "5" = 5),  
       new_column = "ideology5"),
  list(column = "presvote20post", recode_rules = c("1" = 0, "2" = 1),  
       new_column = "presvote_trump_2020"),
  list(column = "presvote16post", recode_rules = c("1" = 0, "2" = 1),  
       new_column = "presvote_trump_2016")
)
cesBYU24 = recodeList(cesBYU24, recode_rules) 
# Only save the newly created ones

new_columns <- names(cesBYU24)
new_columns <- setdiff(new_columns, original_columns)
cesBYU24 <- cesBYU24 %>% 
  select(all_of(new_columns))
```




```{r}
bind_rows(avp23, avp24, wss20, wss24, cesASU, cesBYU) -> electoral_contestation
View(electoral_contestation)
# data
save(electoral_contestation, file = "~/Dropbox/github_repos/electoralContestation/data/electoralConstestation.rda")

```


