---
title: "The Measurement Properties of Electoral Contestation"
header-includes:
    - \usepackage{setspace}\onehalfspacing
author: "Chris Weber"
date: "2025-03-03"
indent: true
output:
  pdf_document: default
---

## Introduction

We begin with a descriptive question: What types of actions do Americans deem acceptable when someone disagrees with the results of an election? Here we focus on levels of support for various actions such as protesting, criticizing election integrity, burning the American flag, ballot recounts, or challenging the outcome in the courts. Moreover, we are interested in understanding the underlying structure of contestation preferences. Is a construct like contestation multidimensional?

On one hand, we might expect that contestation behaviors reside on a single underlying dimension, anchored by a preference for contestation behaviors on one pole, and an opposition to these behaviors at the other pole. However, there are several reasons to expect more nuance, due primarily to different norms surrounding such behaviors. For instance, while recounting ballots and supporting legal means to contest an election are common and generally perceived-to-be acceptable behaviors, attending a march or burning the flag are seen as more active and potentially transgressive behaviors. Contestation behaviors may be effectively disaggregated into forms that pose a relatively high cost for individual citizens (i.e, require action) versus forms of contestation that are passive and impose a low cost for citizens. They can also be viewed as a continuum of behaviors that range from less to more normatively acceptable.

\subsection{Measuring Support for Contestation Behaviors}

We measure support for behaviors aimed at contesting election results with a question battery that captures some of the most prominent ways election results are contested. Respondents were asked,

``Many people are unhappy with the outcomes of elections. How much do you support or oppose each of the following behaviors when people are unhappy with the outcome of an election?''

$\bullet$ Attend a march or demonstration (WSS20[1/2], WSS24, AVPW1, AVPW2, ASU, BYU)

$\bullet$ Burning the American flag (WSS20, WSS24, AVPW1, AVPW2, ASU, BYU)

$\bullet$ Contest the outcome in the courts (WSS20, WSS24, AVPW1, AVPW2, ASU, BYU)

$\bullet$ Support ballot recounts (WSS20, WSS24, AVPW1, AVPW2, ASU, BYU)

$\bullet$ Publicly criticize the integrity or fairness of the election (WSS20[1/2], WSS24, AVPW1, AVPW2, ASU, BYU)

$\bullet$ Engaging in violent acts (WSS24, ASU[Pre], BYU [Post, Control])

$\bullet$ Public officials choosing to certify the election results (AVPW1, AVPW2)

$\bullet$ Refusing to concede an election defeat (AVPW1, AVPW2)

$\bullet$ Call a new election (AVPW1, AVPW2)

$\bullet$ Undermine the election process (WSS24)



Respondents were asked to rate their support for each behavior on a 5-point scale, ranging from 1 (strongly support) to 5 (strongly oppose).

## Recoding and Scaling

We rely on seven data sets in this project. The **Western States Survey** conducted in both 2020 and 2024. The **Arizona Voter Project** election surveys, conducted in 2023 and 2024. And the 2022 Congressional Election Study BYU module and the 2022 Congressional Election Study ASU module.

The $\texttt{electoralContestation}$ package includes a number of helper functions to clean and recode these data. Downloading the package comes with the data $\texttt{electoral\_contestation}$. The total data size is approximately 11,000 respondents.

The data can be downloaded by accessing the package 
```
devtools::install_github("crweber9874\/electoralContestation")
```

```{r, echo = FALSE, message = FALSE, warning = FALSE}
rm(list = ls())
# Install from repository
#devtools::install_github("crweber9874/electoralContestation")
devtools::load_all()
library(dplyr)
library(lavaan)
library(ggplot2)
library(cowplot)
```

You should be able to estimate everything below, by simply the package.


## Models 

In structural equation modeling, full information maximum likelihood (FIML) is frequently used when data are Missing At Random (MAR). FIML involves calculating the likelihood function based on the observed data, then it integrates out the missing values. The likelihood is calculated based on the information available. Insofar as the data are "missing at random" (MAR), FIML can be a useful way in an SEM to handle missing data, and it's not altogether different from the assumptions underlying imputation methods. I am basically using all the information available, across the six studies, using nine contestation items.

Here is how this is done in lavaan. Let's just print out the output for a two factor model, and then we can look at the eigenvalues to see how many factors we should retain.


## Two Factor Model
```{r}
fit <- suppressWarnings(efa(data = electoral_contestation_wide[,c(
                                                       "attend_march",
                                                       "criticize_election",
                                                      # "burn_flag",
                                                       "recount",
                                                       "court",
                                                       "recount",
                                                       "concede_election",
                                                       "state_legislator_certify",
                                                       "violent_protest",
                                                       "new_election",
                                                       "occupy_buildings",
                                                       "object_eligibility",
                                                       "violent_acts"
                                          
                                                       )],
                            nfactors = 2,
                            missing = "fiml",
                           output = "lavaan"
           ))

summary(fit, fit.measures = TRUE, standardized = TRUE)

```

```{r}
fit <- suppressWarnings(efa(data = electoral_contestation_wide[,c(
                                                       "attend_march",
                                                       "criticize_election",
                                                      # "burn_flag",
                                                       "recount",
                                                       "court",
                                                       "recount",
                                                       "concede_election",
                                                       "state_legislator_certify",
                                                       "violent_protest",
                                                       "new_election",
                                                       "occupy_buildings",
                                                       "object_eligibility",
                                                       "violent_acts"
                                          
                                                       )],
                            nfactors = 3 ,
                            missing = "fiml",
                           output = "lavaan"
           ))

summary(fit, fit.measures = TRUE, standardized = TRUE)

```
The two factor model fits well, and the scree plots below supports a two factor solution.

## Factor Structure

```{r}
# \begin{table}[h!] % [h!] tries to place the table "here"
# \centering
# \caption{Electoral Contestation Behaviors} % Add a descriptive caption for your table
# \label{tab:contestation_behaviors} % Give your table a unique label for cross-referencing
# \small
# % Column specification:
# % p{8cm}: First column is a paragraph column, 8cm wide, allowing text to wrap.
# % cc: The next two columns are center-aligned (for numeric data). You could use 'rr' for right-aligned if preferred.
# \begin{tabular}{p{8cm} cc}
#     \toprule % Top horizontal line
#     \textbf{Item} & \textbf{Subversive Contestation} & \textbf{Traditional Contestation} \\
#     \midrule % Mid-table horizontal line (below header)
#     \textit{Attend March} & \textbf{0.570} & 0.050 \\ % 0.050 < 0.3 -> Unbolded
#     \textit{Criticize Election} & \textbf{0.589} & \textbf{0.676} \\
#     \textit{Recounts} & \textbf{0.838} & \textbf{-0.411} \\
#     \textit{Contest in Court} & \textbf{0.763} & 0.167 \\ % 0.167 < 0.3 -> Unbolded
#     \textit{Refusing to Concede the Election} & 0.039 & \textbf{1.131} \\ % 0.039 < 0.3 -> Unbolded
#     \textit{Refuse to Certify} & 0.252 & \textbf{0.740} \\ % 0.252 < 0.3 -> Unbolded
#     \textit{Engage in Violent Protest} & -0.004 & 0.260 \\ % -0.004 < 0.3 (already unbolded), 0.260 < 0.3 -> Unbolded
#     \textit{Call a New Election} & \textbf{0.403} & \textbf{0.535} \\
#     \textit{Occupy Buildings} & -0.047 & \textbf{0.672} \\ % -0.047 < 0.3 (already unbolded)
#     \textit{Object to Voter Eligibility} & 0.186 & \textbf{0.874} \\ % 0.186 < 0.3 (already unbolded)
#     \textit{Engage in Violent Acts} & \textbf{-0.642} & \textbf{1.763} \\
#     \midrule % Changed \hline to \midrule for consistency with booktabs
#     $r_{\text{Subversive, Traditional}}$ & 0.211 & \\ % Added empty cell for alignment
#     \midrule % Changed \hline to \midrule for consistency with booktabs
#     \textbf{Model Fit} & & \\ % Added empty cells for alignment \hline
#     \hline
#     Comparative Fit Index (CFI) & 0.979 & \\ % Added empty cell for alignment
#     Tucker-Lewis Index (TLI) & 0.966 & \\ % Added empty cell for alignment
#     RMSEA & 0.022 & \\ % Added empty cell for alignment
#     \hline
#     N & 14922\\
#     \bottomrule % Bottom horizontal line
# \end{tabular}
# \end{table}
# 
# 
# 
# 
# \begin{table}[h!] % [h!] tries to place the table "here"
# \centering
# \caption{Electoral Contestation Behaviors (One-Factor Model)} % Updated caption
# \label{tab:contestation_behaviors_one_factor} % New label for the one-factor table
# \small
# % Column specification:
# % p{8cm}: First column is a paragraph column, 8cm wide, allowing text to wrap.
# % c: The single data column is center-aligned.
# \begin{tabular}{p{8cm} c}
#     \toprule % Top horizontal line
#     \textbf{Item} & \textbf{Factor Loading} \\ % Updated header for a single factor
#     \midrule % Mid-table horizontal line (below header)
#     \textit{Attend March} & \textbf{0.559} \\ % Updated from f1 attend_march
#     \textit{Criticize Election} & \textbf{0.843} \\ % Updated from f1 criticize_lctn
#     \textit{Recounts} & \textbf{0.500} \\ % Updated from f1 recount
#     \textit{Contest in Court} & \textbf{0.823} \\ % Updated from f1 court
#     \textit{Refusing to Concede the Election} & \textbf{0.639} \\ % Updated from f1 concede_electn
#     \textit{Refuse to Certify} & \textbf{0.643} \\ % Updated from f1 stt_lgsltr_crt
#     \textit{Engage in Violent Protest} & 0.158 \\ % Updated from f1 violent_protst (0.158 < 0.3 -> Unbolded)
#     \textit{Call a New Election} & \textbf{0.681} \\ % Updated from f1 new_election
#     \textit{Occupy Buildings} & \textbf{0.459} \\ % Updated from f1 occupy_buldngs
#     \textit{Object to Voter Eligibility} & \textbf{0.784} \\ % Updated from f1 object_elgblty
#     \textit{Engage in Violent Acts} & 0.203 \\ % Updated from f1 violent_acts (0.203 < 0.3 -> Unbolded)
#     \midrule
#     \textbf{Model Fit} & \\ % Added empty cell for alignment \hline
#     \hline
#     Comparative Fit Index (CFI) & 0.660 \\ % Updated from one-factor model output
#     Tucker-Lewis Index (TLI) & 0.575 \\ % Updated from one-factor model output
#     RMSEA & 0.079 \\ % Updated from one-factor model output
#     \midrule % Consistent midrule before N
#     N & 14922 \\
#     \bottomrule % Bottom horizontal line
# \end{tabular}
# \end{table}
# 
# 
# \begin{table}[h!] % [h!] tries to place the table "here"
# \centering
# \caption{Electoral Contestation Behaviors (Three-Factor Model)} % Updated caption
# \label{tab:contestation_behaviors_three_factor} % New label for the three-factor table
# \small % Keeps the font size small to help fit on the page
# % Column specification:
# % p{8cm}: First column is a paragraph column, 8cm wide, allowing text to wrap.
# % ccc: Three data columns, all center-aligned.
# \begin{tabular}{p{8cm} ccc}
#     \toprule % Top horizontal line
#     \textbf{Item} & \textbf{Traditional} & \textbf{Subversive} & \textbf{Subversive} \\ % Updated headers for three factors
#     \midrule % Mid-table horizontal line (below header)
#     \textit{Attend March} & \textbf{1.284} & \textbf{-0.852} & 0.000 \\
#     \textit{Criticize Election} & \textbf{0.646} & \textbf{0.443} & 0.014 \\
#     \textit{Recounts} & \textbf{0.586} & -0.001 & \textbf{-0.961} \\
#     \textit{Contest in Court} & \textbf{0.758} & 0.024 & -0.288 \\
#     \textit{Refusing to Concede the Election} & 0.012 & \textbf{0.923} & \textbf{0.373} \\
#     \textit{Refuse to Certify} & 0.171 & \textbf{0.666} & 0.110 \\
#     \textit{Engage in Violent Protest} & 0.092 & 0.112 & 0.152 \\
#     \textit{Call a New Election} & 0.240 & \textbf{0.568} & -0.056 \\
#     \textit{Occupy Buildings} & -0.006 & \textbf{0.513} & 0.297 \\
#     \textit{Object to Voter Eligibility} & -0.098 & \textbf{0.998} & -0.020 \\
#     \textit{Engage in Violent Acts} & 0.002 & \textbf{0.577} & \textbf{1.208} \\
#     \midrule
#     \multicolumn{4}{l}{\textbf{Factor Correlations}} \\ % Heading for correlations, spanning all 4 columns
#     \addlinespace[0.5em] % Small vertical space for readability
#     $r_{F1, F2}$ & 0.715 & & \\ % Correlation between Factor 1 and Factor 2
#     $r_{F1, F3}$ & 0.210 & & \\ % Correlation between Factor 1 and Factor 3
#     \midrule
#     \multicolumn{4}{l}{\textbf{Model Fit}} \\ % Heading for model fit, spanning all 4 columns
#     \hline
#     Comparative Fit Index (CFI) & 0.998 & & \\ % Updated from three-factor model output
#     Tucker-Lewis Index (TLI) & 0.997 & & \\ % Updated from three-factor model output
#     RMSEA & 0.007 & & \\ % Updated from three-factor model output
#     \midrule
#     N & 14922 & & \\ % Number of observations
#     \bottomrule % Bottom horizontal line
# \end{tabular}
# \end{table}
```


```{r}
fit <- suppressWarnings(efa(data = electoral_contestation_wide[,c(
                                                       "attend_march",
                                                       "criticize_election",
                                                       #"burn_flag",
                                                       "recount",
                                                       "court",
                                                       "recount",
                                                       "concede_election",
                                                       "state_legislator_certify",
                                                       "violent_protest",
                                                       "new_election",
                                                       "occupy_buildings",
                                                       "object_eligibility",
                                                       "violent_acts"
                                          
                                                       )],
                            nfactors = 1,
                            missing = "fiml",
                           output = "lavaan"
           ))

summary(fit, fit.measures = TRUE, standardized = TRUE)


cov_matrix = lavInspect(fit, "implied")$cov
sd_vec <- sqrt(diag(cov_matrix))
cor_matrix <- cov_matrix / (sd_vec %*% t(sd_vec))

# Eigenvalues of the correlation matrix
eigenvalues <- eigen(cor_matrix)$values
eigenvalues
eigen_df <- data.frame(
  Factor = 1:length(eigenvalues),
  Eigenvalue = eigenvalues
)


# Scree plot using ggplot2 (dots connected by lines)
scree_plot <- ggplot(eigen_df, aes(x = Factor, y = Eigenvalue)) +
  geom_line(alpha = 0.6) +  # Connect the dots with lines
  geom_point(size = 6, alpha = 0.3, color = "black") + # Add dots at each point
  labs(
    title = "Electoral Contestation\nScree Plot",
    x = "Factors",
    y = "Eigenvalue"
  ) +
  theme_minimal() + 
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.line = element_line(colour = "black"),
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 14),
    plot.title = element_text(size = 16)
  ) + 
  geom_hline(yintercept = 1, linetype = "dashed", color = "darkgrey")

scree_plot
ggsave("scree_plot.png", width = 8, height = 8, dpi = 500)

```


```{r}
summary(fit, fit.measures = TRUE, standardized = TRUE)

```

The two factor makes sense here -- both eigenvalues exceed the (arbitrary) cut point and the "elbow" is also at 2. The fit is pretty terrible for the one-factor model. Here's the fit for a confirmatory factor model, following this structure

```{r}
library(dplyr)
library(lavaan)
library(tidyr)
  ggtheme =
    theme(
      plot.title =  ggplot2::element_text(face = "bold", hjust = 0, vjust = 0, colour = "#3C3C3C", size = 12),
      axis.text.x = ggplot2::element_text(size = 12, colour = "#535353"),
      axis.text.y = ggplot2::element_text(size = 12, colour = "#535353"),
      axis.title =  ggplot2::element_text(size = 12, colour = "#535353"),
      axis.title.y = ggplot2::element_text(size = 12, colour = "#535353", vjust = 1.5),
      # Change facet
#      strip.background = ggplot2::element_rect(fill = "#D0D0D0", colour = "#D0D0D0"),
      axis.ticks = ggplot2::element_blank(),
      panel.grid.major = ggplot2::element_line(colour = "#D0D0D0", size = .25),
      panel.background = ggplot2::element_rect(fill = "white"),
      legend.text = ggplot2::element_text(size = 10),
      legend.title = ggplot2::element_text(size = 10),
      # chnge facet wrap label size
      )



  # & {Subversive Contestation} & {Traditional Contestation} \\
  # \midrule
  # \textit{Engaging in violent acts} & -0.253 & \textbf{0.630 }\\
  # \textit{Refusing to Concede the Election} & 0.026 & \textbf{0.885} \\
  # \textit{Publicly criticize the integrity or fairness of the election} & \textbf{0.389 }&\textbf{ 0.649} \\
  # \textit{Attend a march or demonstration} & \textbf{0.549 }& -0.132 \\
  # \textit{Contest the outcome in the courts} & \textbf{0.798} & 0.129 \\
  # \textit{Support election recounts} &\textbf{ 0.590} & 0.010 \\
  # \textit{Call a new election} & 0.215 &\textbf{ 0.674} \\
  # \textit{Undermine the election process} & 0.147 & \textbf{0.700} \\
  # \textit{State officials refusing to certify the election results} & -0.004 & \textbf{0.872} \\
  # \textit{Occupy} & -0.168 & 0.741 \\
  # 


electoral_contestation_wide %>%  
  select("attend_march",
         "criticize_election",
         "recount",
         "court",
         "recount",
         "concede_election",
         "state_legislator_certify",
         "violent_protest",
         "new_election",
         "occupy_buildings",
         "object_eligibility",
         "violent_acts") %>%
   rename(`Refuse to Certify` = state_legislator_certify,
          `Refuse to Concede` = concede_election,
          `Criticize Election` = criticize_election,
          `Attend a\nMarch or Rally` = attend_march,
          `Court` = court,
          `Recounts` = recount,
          `Call a New Election` = new_election,
          `Object to Eligibility` = object_eligibility,
          `Occupy Buildings` = occupy_buildings,
          `Violent Acts` = violent_acts,
          `Violent Protest` = violent_protest) %>%
    pivot_longer(cols = everything(), names_to = "variable", values_to = "value")  %>%
  ggplot(aes(x = value)) +
  ggtheme + 
    facet_wrap(~variable, ncol = 3, scales = "free") + 
  geom_histogram(aes(y = ..density..), binwidth = 1, 
                 fill = "grey", color = "darkgrey",
                 alpha = 0.3) +
  geom_density(color = "darkblue", bw = 0.5) +
 theme(strip.text = element_text(size = 11),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()
) + 
  labs(title = "Electoral Contestation\nDistribution of Responses",
       x = "Disagree-Agree",
       y = "Density")
ggsave("electoral_contestation_histograms.png", width = 8, height = 8, dpi = 500)
```


```{r}
library(vayr)
library(estimatr)

electoral_contestation_long %>%
  mutate(contestation_clean = case_when(
    contestation == "attend_march" ~ "Attend March",
    contestation == "criticize_election" ~ "Criticize Election",
    contestation == "recount" ~ "Ballot Recounts",
    contestation == "court" ~ "Contest in Court",
    contestation == "concede_election" ~ "Refuse to Concede",
    contestation == "state_legislator_certify" ~ "Refuse to Certify",
    contestation == "violent_protest" ~ "Violent Protest",
    contestation == "new_election" ~ "Call a New Election",
    contestation == "occupy_buildings" ~ "Occupy Buildings",
    contestation == "object_eligibility" ~ "Object to Eligibility",
    contestation == "violent_acts" ~ "Violent Acts" ,
    TRUE ~ NA_character_
  )) %>% filter(!is.na(contestation_clean))  -> electoral_contestation_long

summary_df <- electoral_contestation_long |>
  select(condition,
         contestation_clean,
         survey,
         contestation_value) |>
  group_by(survey, contestation_clean) |>
  reframe(tidy(lm_robust(contestation_value ~ 1))) 

# A df for direct labels
label_df <- summary_df 


  # filter(!is.na(party_identification3)) |>
  # mutate(
  #   electoral_contestation= case_when(
  #     party_identification3 == 1 ~ conf.low - 0.15,
  #     party_identification3 == 2 ~ conf.low - 0.15,
  #     party_identification3 == 3 ~ conf.high + 0.15
  #   ),
  #   party_identification3 = factor(party_identification3, 
  #                                         levels = c(1, 2, 3),
  #                                         labels = c("Democrat", "Independent", "Republican"))
  # )
  # 
ggplot(electoral_contestation_long, aes(survey, contestation_value, group = contestation_clean)) +
      facet_wrap(~contestation_clean, nrow = 6)+ 
  # the data
  geom_point(position = position_circlepack(density = 6,
                                                aspect_ratio = 0.5),
             size = 0.001, alpha = 0.3, color = "grey") +
  # the statistical model
  geom_line(data = summary_df, aes(x = survey, y = estimate),  
            position = position_dodge(width = 0.5), linewidth = 0.5, color= "black") +  
  geom_point(data = summary_df, aes(x = survey, y = estimate),  
             position = position_dodge(width = 0.5), size = 3, color = "black") +
  # geom_linerange(data = summary_df, aes(x = survey, y = estimate,
  #                                       ymin = conf.low, ymax = conf.high, color = "darkgrey"),
  #                position = position_dodge(width = 0.5)) + 
     theme_bw() +
     theme(legend.position = "none",
           strip.background = element_blank(),
           panel.grid.minor = element_blank()) +
  labs(title = "Electoral Contestation\nSupport for Contestation Behaviors",
       x = "Survey",
       y = "Contestation Behaviors (1: Strongly Oppose to\n5: Strongly Support)") +
  scale_y_continuous(breaks = seq(1, 5, 1), limits = c(0, 6)) +
  # rotate x axis labels 45
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) 

ggsave("electoral_contestation_vayr.png", width = 8, height = 8, dpi = 500)


```


```{r}

library(reshape2)
cov_matrix = lavInspect(fit, "implied")$cov
sd_vec <- sqrt(diag(cov_matrix))
correlation_matrix <- cov_matrix / (sd_vec %*% t(sd_vec))


# Reorder the correlation matrix using hierarchical clustering
hc <- hclust(dist(correlation_matrix))
reordered_vars <- rownames(correlation_matrix)[hc$order]
reordered_matrix <- correlation_matrix[reordered_vars, reordered_vars]

correlation_long <- melt(reordered_matrix)

# Create the heatmap
heatmap_plot <- ggplot(data = correlation_long, aes(x = Var1, y = Var2, fill = value)) +
  geom_tile() +
  scale_fill_gradient2(low = "white", mid = "white", high = "darkgrey", 
                       midpoint = 0, limit = c(-1, 1), space = "Lab", 
                       name="Pearson\nCorrelation") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, 
                                   size = 10, hjust = 1)) +
  coord_fixed() + 
  labs(title = "Electoral Contestation") + 
  xlab("Contestation Items") +
  ylab("Contestation Items")

heatmap_plot
ggsave("corr.png", width = 4, height = 4)
```









```{r}
library(dplyr)
library(lavaan)

# electoral_contestation %>%
#   mutate(year = ifelse(survey == "wss20", 2020,
#                        ifelse(survey == "wss24", 2024,
#                               ifelse(survey == "avpw1", 2023,
#                                      ifelse(survey == "avpw2", 2024,
#                                             ifelse(survey == "cesBYU", 2022,
#                                                    ifelse(survey == "cesASU", 2022,
#                                                           NA)))))),  
#     post2022 = abs(as.numeric(year > 2022)),
#          authoritarianism = rowMeans(electoral_contestation[, c("auth_1", "auth_2",
#                                                                  "auth_3", "auth_4")], na.rm = TRUE),
#          trump_voter = ifelse(presvote_trump_2020 == 1, 1, 0),
#          trump_voter = ifelse(is.na(trump_voter), presvote_trump_2024, trump_voter),
#          trumpXauthoritarianism = trump_voter * authoritarianism
#   ) -> dat

model <- "f1 =~ attend_march + criticize_election + recount + court 
          f2 =~ state_legislator_certify + concede_election + criticize_election + recount + 
          new_election + occupy_buildings + object_eligibility + violent_acts + violent_protest
          "

# model = paste(model, 
#               "f1 ~", "trump_voter", "\n",
#               "f2 ~", "trump_voter"
#               )

# Perform multiple group CFA with measurement invariance constraints
fit_multi <- suppressWarnings(sem(model,
                                 data = electoral_contestation_wide,
                                 # group = "trump_voter",
                                 # group.equal = c("loadings", "intercepts", "residuals"),
                                 missing = "fiml"))
summary(fit_multi, fit.measures = TRUE, standardized = TRUE)
# # This is somewhat clunky with fiml
 factor_scores <- lavPredict(fit_multi, method = "regression")
 factor_scores_df <- as.data.frame(factor_scores)
 factor_scores_df$row_id <- as.integer(rownames(factor_scores_df))
 electoral_contestation_wide$row_id <- 1:nrow(electoral_contestation_wide)
 electoral_contestation_with_scores <- left_join(electoral_contestation_wide, factor_scores_df, by = "row_id")
 electoral_contestation_fscors <- electoral_contestation_with_scores %>% select(-row_id)
```

```{r}
# estimate a BRMS Model regressing y1 and y2 correlated on trump_voter
library(brms)
# Create a new data frame with the factor scores
electoral_contestation_fscors %>%
  mutate(partisan = ifelse(party_identification3 != 2  & !is.na(party_identification3), "Partisan", "Independent"),
         ideology = case_when(
           ideology5 == 5 ~ "Conservative",
           ideology5 == 4 ~ "Conservative",
           ideology5 == 3 ~ "Moderate",
           ideology5 == 2 ~ "Liberal",
           ideology5 == 1 ~ "Liberal",
            TRUE ~ NA
         ),
         extreme = ifelse(ideology5 == 1 | ideology5 == 5, 1, 0),
         ) -> dat
    
  
# Fit the model
fit_brms <- brm(
    bf(f2 ~ black + latino + white + faminc  + ideology   + as.factor(party_identification3)) +
    bf(f1 ~ black + latino + white +  faminc + ideology + as.factor(party_identification3)),  
  data = dat,
  family = gaussian(),
  iter = 2000,
  chains = 4,
  cores = 10,
)
summary(fit_brms)
```
```{r}
library(modelr)
library(tidybayes)
library(haven)
library(reshape2)

electoral_contestation %>% 
    mutate(post2020 = ifelse(year>2022, 1, 0),
           authoritarianism = rowMeans(electoral_contestation[, 
                                                            c("auth_1", "auth_2",
                                                              "auth_3", "auth_4")], na.rm = TRUE),
         trump_voter = ifelse(presvote_trump_2020 == 1, 1, 0),
         trump_voter = ifelse(is.na(trump_voter), 
                              presvote_trump_2024, trump_voter),
         trumpXauthoritarianism = trump_voter * authoritarianism) %>%
  data_grid(
        female  = mean(female,  na.rm = TRUE), 
        college = mean(college, na.rm = TRUE), 
        faminc  = mean(faminc,  na.rm = TRUE),
        black   = mean(black,   na.rm = TRUE), 
        latino  = mean(latino,  na.rm = TRUE), 
        white   = mean(white,   na.rm = TRUE), 
        authoritarianism = seq_range(authoritarianism, n = 2),
        trump_voter = c(0,1),
        post2022    = c(0,1))   %>%
      add_epred_draws(fit_brms) %>% 
      group_by(trump_voter, post2022, .category)  %>%
  summarize(
    factor_score = mean(.epred),
    .lower = quantile(.epred, 0.025),
    .upper = quantile(.epred, 0.975)
  ) %>%
  mutate(
    trump_voter = ifelse(trump_voter == 1, "Trump Voter", "Non-Trump Voter"),
    post2022 = ifelse(post2022 == 1, "2023-2024", "2020-2022"),
    factor = ifelse(.category == "f1", "Subversive", "Traditional"))  %>%
   ggplot(aes(x = as.factor(post2022),
                  y = factor_score, 
                  ymin=.lower, 
                  ymax=.upper,
                  group = factor,
                  color = factor
             ))+
facet_wrap(~ as.factor(trump_voter) ) +
  geom_point(size=4, alpha = 0.3, position = position_dodge(width = 0.2)) +
  geom_errorbar(width = 0.10, alpha=0.75, colour = "black", position = position_dodge(width = 0.2)) +
  ggtitle("Electoral Contestation\nBy Voting and Survey Year") +
  scale_y_continuous("Factor Score")+
  scale_x_discrete("Survey") +
  theme_minimal() +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.line = element_line(colour = "black"),
    axis.text = element_text(size = 10),
    axis.title = element_text(size = 12),
    plot.title = element_text(size = 14)
  )+
  # legend
  theme(legend.position = "bottom",
        legend.title = element_blank(),
        legend.text = element_text(size = 10),
        legend.key.size = unit(0.5, "cm"),
        legend.key.width = unit(1, "cm")) +
  # change colors
  scale_color_manual(values = c("Subversive" = "black", "Traditional" = "grey")) 

ggsave("electoral_contestation_brms_trump.png", width = 8, height = 6, dpi = 500)

```




```{r}
# estimate a BRMS Model regressing y1 and y2 correlated on trump_voter
library(brms)
# Create a new data frame with the factor scores
electoral_contestation_fscors %>%
  mutate(post2022 = ifelse(year>2022, 1, 0),
         authoritarianism = rowMeans(electoral_contestation[, c("auth_1", "auth_2",
                                                                 "auth_3", "auth_4")], 
                                     na.rm = TRUE),
         trump_voter = ifelse(presvote_trump_2020 == 1, 1, 0),
         trump_voter = ifelse(is.na(trump_voter), presvote_trump_2024, trump_voter),
         trumpXauthoritarianism = trump_voter * authoritarianism,
         trump_voter = as.factor(trump_voter),
         pid = zero.one(party_identification7),
         extremity = zero.one(abs(4-party_identification7)),
         conservative = ifelse(ideology5 == 4 | ideology5 == 5, 1, 0),
         liberal = ifelse(ideology5 == 1 | ideology5 == 2, 1, 0),
         moderate = ifelse(ideology5 == 4 , 1, 0)) -> dat
  
  
# Fit the model
fit_brms <- brm(
    bf(f2 ~ trump_voter + black + latino + white + college + faminc + female + authoritarianism + post2022 + trump_voter:post2022) + 
    bf(f1 ~ trump_voter + black + latino + white + college + faminc + female + authoritarianism + post2022 + trump_voter:post2022),  
  data = dat,
  family = gaussian(),
  iter = 2000,
  chains = 4,
  cores = 10,
)
summary(fit_brms)
```


```{r}
# Fit the model
fit_brms <- brm(
    bf(f2 ~ extremity + pid + black + latino + white + college + faminc + female + 
            authoritarianism + post2022 + pid:post2022 + extremity:post2022) + 
    bf(f1 ~ extremity + pid + black + latino + white + college + faminc + female + 
            authoritarianism + post2022 + pid:post2022 + extremity:post2022),
  data = dat,
  family = gaussian(),
  iter = 2000,
  chains = 4,
  cores = 10,
)
```

```{r}
library(modelr)
library(tidybayes)
library(haven)
library(reshape2)

electoral_contestation %>% 
  mutate(post2022 = ifelse(year>2022, 1, 0),
         authoritarianism = rowMeans(electoral_contestation[, c("auth_1", "auth_2",
                                                                 "auth_3", "auth_4")], 
                                     na.rm = TRUE),
         trump_voter = ifelse(presvote_trump_2020 == 1, 1, 0),
         trump_voter = ifelse(is.na(trump_voter), presvote_trump_2024, trump_voter),
         trumpXauthoritarianism = trump_voter * authoritarianism,
         trump_voter = as.factor(trump_voter),
         pid = zero.one(party_identification7),
         extremity = zero.one(abs(4-party_identification7)),
         conservative = ifelse(ideology5 == 4 | ideology5 == 5, 1, 0),
         liberal = ifelse(ideology5 == 1 | ideology5 == 2, 1, 0),
         moderate = ifelse(ideology5 == 4 , 1, 0)) %>%
data_grid(
        female  = mean(female,  na.rm = TRUE), 
        college = mean(college, na.rm = TRUE), 
        faminc  = mean(faminc,  na.rm = TRUE),
        black   = mean(black,   na.rm = TRUE), 
        latino  = mean(latino,  na.rm = TRUE), 
        white   = mean(white,   na.rm = TRUE), 
        authoritarianism = seq_range(authoritarianism, n = 2),
        pid = seq_range(pid, n = 7),
        post2022    = c(0,1),
        extremity = mean(extremity, na.rm = TRUE))   %>%
      add_epred_draws(fit_brms) %>% 
      group_by(pid, post2022, .category)  %>%
  summarize(
    factor_score = mean(.epred),
    .lower = quantile(.epred, 0.025),
    .upper = quantile(.epred, 0.975)
  ) %>%
  mutate(
    post2022 = ifelse(post2022 == 1, "2023-2024", "2020-2022"),
    factor = ifelse(.category == "f1", "Subversive", "Traditional")
    )  %>%
   ggplot(aes(x = as.numeric(pid),
                  y = factor_score, 
                  ymin=.lower, 
                  ymax=.upper,
                  group = factor,
                  color = factor
             ))+
facet_wrap(~ as.factor(post2022) ) +
  geom_point(size=4, alpha = 0.3) +
  geom_errorbar(width = 0.01, alpha=0.75, colour = "black") +
  geom_line() + 
  ggtitle("Electoral Contestation\nBy Party Identification and Survey Year") +
  scale_y_continuous("Factor Score")+
  scale_x_continuous("Party Identification [0 = Strong Democrat, 1 = Strong Republican]") +
  theme_minimal() +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.line = element_line(colour = "black"),
    axis.text = element_text(size = 10),
    axis.title = element_text(size = 12),
    plot.title = element_text(size = 14)
  )+
  # legend
  theme(legend.position = "bottom",
        legend.title = element_blank(),
        legend.text = element_text(size = 10),
        legend.key.size = unit(0.5, "cm"),
        legend.key.width = unit(1, "cm")) +
  scale_color_manual(values = c("Subversive" = "black", "Traditional" = "grey")) 
  #
ggsave("electoral_contestation_brms_pid.png", width = 8, height = 6, dpi = 500)
```


```{r}
library(modelr)
library(tidybayes)
library(haven)
library(reshape2)

electoral_contestation %>% 
  mutate(post2022 = ifelse(year>2022, 1, 0),
         authoritarianism = rowMeans(electoral_contestation[, c("auth_1", "auth_2",
                                                                 "auth_3", "auth_4")], 
                                     na.rm = TRUE),
         trump_voter = ifelse(presvote_trump_2020 == 1, 1, 0),
         trump_voter = ifelse(is.na(trump_voter), presvote_trump_2024, trump_voter),
         trumpXauthoritarianism = trump_voter * authoritarianism,
         trump_voter = as.factor(trump_voter),
         pid = zero.one(party_identification7),
         extremity = zero.one(abs(4-party_identification7)),
         conservative = ifelse(ideology5 == 4 | ideology5 == 5, 1, 0),
         liberal = ifelse(ideology5 == 1 | ideology5 == 2, 1, 0),
         moderate = ifelse(ideology5 == 4 , 1, 0)) %>%
data_grid(
        female  = mean(female,  na.rm = TRUE), 
        college = mean(college, na.rm = TRUE), 
        faminc  = mean(faminc,  na.rm = TRUE),
        black   = mean(black,   na.rm = TRUE), 
        latino  = mean(latino,  na.rm = TRUE), 
        white   = mean(white,   na.rm = TRUE), 
        authoritarianism = seq_range(authoritarianism, n = 2),
        pid = mean(pid, na.rm = TRUE),
        post2022    = c(0,1),
        extremity = seq_range(extremity, n = 5))   %>%
      add_epred_draws(fit_brms) %>% 
      group_by(extremity, post2022, .category)  %>%
  summarize(
    factor_score = mean(.epred),
    .lower = quantile(.epred, 0.025),
    .upper = quantile(.epred, 0.975)
  ) %>%
  mutate(
    post2022 = ifelse(post2022 == 1, "2023-2024", "2020-2022"),
    factor = ifelse(.category == "f1", "Subversive", "Traditional")
    )  %>%
   ggplot(aes(x = as.numeric(extremity),
                  y = factor_score, 
                  ymin=.lower, 
                  ymax=.upper,
                  group = factor,
                  color = factor
             ))+
facet_wrap(~ as.factor(post2022) ) +
  geom_point(size=4, alpha = 0.3) +
  geom_errorbar(width = 0.01, alpha=0.75, colour = "black") +
  geom_line() + 
  ggtitle("Electoral Contestation\nBy Party Extremity and Survey Year") +
  scale_y_continuous("Factor Score")+
  scale_x_continuous("Party Extremity [0 =  Moderate, 1 = Extreme]") +
  theme_minimal() +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.line = element_line(colour = "black"),
    axis.text = element_text(size = 10),
    axis.title = element_text(size = 12),
    plot.title = element_text(size = 14)
  )+
  # legend
  theme(legend.position = "bottom",
        legend.title = element_blank(),
        legend.text = element_text(size = 10),
        legend.key.size = unit(0.5, "cm"),
        legend.key.width = unit(1, "cm")) +
  scale_color_manual(values = c("Subversive" = "black", "Traditional" = "grey")) 
  #
ggsave("electoral_contestation_brms_extremity.png", width = 8, height = 6, dpi = 500)
```



```{r}
library(electoralContestation)
library(dplyr)
library(lavaan)
library(ggplot2)
library(cowplot)

fit <- suppressWarnings(efa(data = electoral_contestation[,c(
                                                       "violent",
                                                       "concede",
                                                       "criticize_election",
                                                       "attend_march",
                                                       "court",
                                                       "recount",
                                                       "new_election",
                                                       "undermine",
                                                       "state_certify",
                                                       "certify_occupy",
                                                       "voteconfidence",
                                                       "free_fair"
                                                       )],
                            nfactors = 3,
                            missing = "fiml",
                           output = "lavaan"
           ))

summary(fit, fit.measures = TRUE)
```

library(dplyr)
library(lavaan)

# Assuming 'dat' and 'model' are already defined

# Configural Invariance (baseline)
fit_configural <- suppressWarnings(sem(model,
                                      data = dat,
                                      group = "trump_voter",
                                      missing = "fiml"))

# Metric Invariance (loadings constrained)
fit_metric <- suppressWarnings(sem(model,
                                  data = dat,
                                  group = "trump_voter",
                                  group.equal = c("loadings"),
                                  missing = "fiml"))

# Scalar Invariance (loadings and intercepts constrained)
fit_scalar <- suppressWarnings(sem(model,
                                  data = dat,
                                  group = "trump_voter",
                                  group.equal = c("loadings", "intercepts"),
                                  missing = "fiml"))

# Strict Invariance (loadings, intercepts, and residuals constrained)
fit_strict <- suppressWarnings(sem(model,
                                  data = dat,
                                  group = "trump_voter",
                                  group.equal = c("loadings", "intercepts", "residuals"),
                                  missing = "fiml"))

# Partial Metric Invariance (constrain all but 1 or 2 loadings)
partial_metric_1 <- suppressWarnings(sem(model,
                                          data = dat,
                                          group = "trump_voter",
                                          group.equal = c("loadings"),
                                          group.partial = list(list("f1 =~ violent")), # Free violent loading
                                          missing = "fiml"))

partial_metric_2 <- suppressWarnings(sem(model,
                                          data = dat,
                                          group = "trump_voter",
                                          group.equal = c("loadings"),
                                          group.partial = list(list("f1 =~ violent", "f2 =~ attend_march")), # Free violent and attend_march loadings
                                          missing = "fiml"))

# Partial Scalar Invariance (constrain all but 1 or 2 intercepts)
partial_scalar_1 <- suppressWarnings(sem(model,
                                          data = dat,
                                          group = "trump_voter",
                                          group.equal = c("loadings", "intercepts"),
                                          group.partial = list(list("violent~1")), # Free violent intercept
                                          missing = "fiml"))

partial_scalar_2 <- suppressWarnings(sem(model,
                                          data = dat,
                                          group = "trump_voter",
                                          group.equal = c("loadings", "intercepts"),
                                          group.partial = list(list("violent~1", "attend_march~1")), # Free violent and attend_march intercepts
                                          missing = "fiml"))

# Compare models
anova_config_metric <- anova(fit_configural, fit_metric)
anova_metric_scalar <- anova(fit_metric, fit_scalar)
anova_scalar_strict <- anova(fit_scalar, fit_strict)
anova_metric_partial1 <- anova(fit_metric, partial_metric_1)
anova_metric_partial2 <- anova(fit_metric, partial_metric_2)
anova_scalar_partial1 <- anova(fit_scalar, partial_scalar_1)
anova_scalar_partial2 <- anova(fit_scalar, partial_scalar_2)

# Print results
print(anova_config_metric)
print(anova_metric_scalar)
print(anova_scalar_strict)
print(anova_metric_partial1)
print(anova_metric_partial2)
print(anova_scalar_partial1)
print(anova_scalar_partial2)

# Summarize model fits
summary(fit_configural, fit.measures = TRUE)
summary(fit_metric, fit.measures = TRUE)
summary(fit_scalar, fit.measures = TRUE)
summary(fit_strict, fit.measures = TRUE)
summary(partial_metric_1, fit.measures = TRUE)
summary(partial_metric_2, fit.measures = TRUE)
summary(partial_scalar_1, fit.measures = TRUE)
summary(partial_scalar_2, fit.measures = TRUE)

# Compare model fits using anova()
anova_config_metric <- anova(fit_configural, fit_metric)
anova_metric_scalar <- anova(fit_metric, fit_scalar)
anova_scalar_strict <- anova(fit_scalar, fit_strict)

# Print results
print(anova_config_metric)
print(anova_metric_scalar)
print(anova_scalar_strict)

# # Also look at the model fit statistics
# summary(fit_configural, fit.measures = TRUE)
# summary(fit_metric, fit.measures = TRUE)
# summary(fit_scalar, fit.measures = TRUE)
# summary(fit_strict, fit.measures = TRUE)
```



electoral_contestation %>%
  mutate(post2020 = abs(as.numeric(year > 2020)),
         authoritarianism = rowMeans(electoral_contestation[, c("auth_1", "auth_2",
                                                               "auth_3", "auth_4")], na.rm = TRUE),
         trump_voter =  ifelse(presvote_trump_2020 == 1, 
                              1 , 0 ),
         trump_voter =  ifelse(is.na(trump_voter), presvote_trump_2024, trump_voter),
        trumpXauthoritarianism = trump_voter * authoritarianism
           )  -> dat


# Assuming 'electoral_contestation', 'year', 'party_identification3', and 'model' are defined
model = "f1 =~ violent + state_certify + concede +  criticize_election + new_election + undermine  + certify_occupy\n    
f2 =~  attend_march + court + recount\n"
# model = paste(model, 
#               "f1 ~", "trump_voter", "\n",
#               "f2 ~", "trump_voter"
#               )

# Perform multiple group CFA with measurement invariance constraints
fit_multi <- suppressWarnings(sem(model,
                                 data = dat,
                                 group = "trump_voter",
                                group.equal = c("loadings", "intercepts", "residuals"),
                                 missing = "fiml"))

# Compare different fits




# extract factor scores and append to data
dat <- cbind(dat, 
             lavPredict(fit_multi, type = "lv") %>% as.data.frame() %>% 
               rename(f1 = f1, f2 = f2))


# Extract parameter estimates
parameter_estimates <- parameterEstimates(fit_multi) %>% 
  filter((lhs == "f1"| lhs =="f2") & op == "~" & rhs == "trump_voter") -> mean.estimates
  

# Fitted values in lavaan
lavPredict(fit_multi,
             newdata =  dat %>% mutate(trump_voter = 1),
             type = "lv") %>%
  as.data.frame() %>%
  summarise(f1 = mean(f1),
            f2 = mean(f2)) -> fitted_values



```

The fit is quite good. 

Let's now look at the eigenvalues *within* survey wave.

### Within Each Survey Wave

First, I created a simple function to prevent recycling the same code.

```{r}
runEFA = function(factors = 2, 
                  data = electoral_contestation){
  
         df = data # Change if need be
         fit <- suppressWarnings(
                            efa(data = data,
                            nfactors = factors,
                            missing = "fiml"
           ))
       if(factors == 2) eigenvalues = summary(fit)[[10]]$nf2[[1]]$eigvals[[1]]
       if(factors == 3) eigenvalues =summary(fit)[[10]]$nf3[[1]]$eigvals[[1]]
       if(factors == 4) eigenvalues =summary(fit)[[10]]$nf4[[1]]$eigvals[[1]]
    
         eigen_df <- data.frame(
           Factor = 1:length(eigenvalues),
           Eigenvalue = eigenvalues
         )                                      
scree_plot <- ggplot(eigen_df, aes(x = Factor, y = Eigenvalue)) +
  geom_line(alpha = 0.6) +  
  geom_point(size = 3, alpha = 0.3, color = "black") + 
  labs(
    title = "Electoral Contestation\nScree Plot",
    x = "Factors",
    y = "Eigenvalue"
  ) +
  theme_minimal() + 
  # style a hell of lot better
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.line = element_line(colour = "black"),
    axis.text = element_text(size = 10),
    axis.title = element_text(size = 12),
    plot.title = element_text(size = 14)
  ) + 
  geom_hline(yintercept = 1, linetype = "dashed", color = "darkgrey")

         
         return(list(fit = fit, 
                     eigen_df = eigen_df, 
                     scree_plot = scree_plot))
}
electoral_contestation %>%
  select(c("state_certify", "concede","criticize_election", "attend_march", "court",  "recount", "new_election", "undermine", "violent") ) %>%
runEFA(factors = 2)
```

```{r}
options(scipen = 999, repr.plot.width = 12, repr.plot.height = 10)

var = c("state_certify", "concede","criticize_election", "attend_march", "court",  "recount", "new_election", "undermine", "violent", "survey") 
results <- list()
for (s in unique(electoral_contestation$survey)) {
        results[[s]] <-
            electoral_contestation %>%
            filter(survey == s) %>%
            select(all_of(var)) %>%
            select(!survey) %>%
              select_if(~ !all(is.na(.)))  %>%
          na.omit()

}
        
avpw1<- runEFA(factors = 2, data = results[[1]])[[3]] +
  ggplot2::ggtitle(paste("Arizona Voter Project\n(2023)"))
avpw2<- runEFA(factors = 2, data = results[[2]])[[3]] +
  ggplot2::ggtitle(paste("Arizona Voter Project\n(2024)"))
wss20<- runEFA(factors = 1, data = results[[3]])[[3]] +
  ggplot2::ggtitle(paste("Western States Survey\n(2020)"))
wss24<- runEFA(factors = 1, data = results[[4]])[[3]] +
  ggplot2::ggtitle(paste("Western States Survey\n(2024)"))
cesASU<- runEFA(factors = 1, data = results[[5]])[[3]] +
  ggplot2::ggtitle(paste("ASU CES\n(2022)"))

scree_plots <- cowplot::plot_grid(avpw1, avpw2, wss20, wss24, cesASU, ncol = 3)
scree_plots

```
These surveys have differing numbers of items, but the factor structure -- 2 -- looks pretty similar across surveys. There seems to be "conventional/traditional" factor and a "socially transgressive" factor -- newer means of electoral contestation, like "calling a new election."

Here is the correlation matrix across surveys.
```{r}
library(dplyr)


```

Note how items like "undermine" and "violent" correlated with "new election" but not with "court" or "recount." 

## Party Differences

The groups are 1 = Democrat, 2 = Independent, 3 = Republican. Republicans score highest on both forms of contestation.

```{r}
library(lavaan)
# Multiple Group Model
model <- ' f1 =~ violent + state_certify + concede +  criticize_election + new_election + undermine
           f2 =~  attend_march + court + recount'

suppressWarnings(cfa(model, 
                            data = electoral_contestation,
                            missing = "fiml",
                            group = "party_identification3",
#                           group.equal = c("loadings", "intercepts", "residuals")
           )) %>% summary(fit.measures = TRUE)

```
This is the same thing.

```{r}
electoral_contestation %>%
        mutate(
           republican = as.numeric(party_identification3 == 3),
           independent = as.numeric(party_identification3 == 2),
           democrat = as.numeric(party_identification3 == 1),
        ) -> df


model <- ' f1 =~ violent + state_certify + concede +  
                 criticize_election + new_election + undermine
           f2 =~  attend_march + court + recount
           f1 ~ independent + democrat 
           f2 ~ independent + democrat'

suppressWarnings(sem(model, 
                            data = df,
                            missing = "fiml",
           )) %>% summary(fit.measures = TRUE)

```
```


## Appendix
This isn't pretty. We'd need to clean this up if we decide to present this.


```{r}
vars = c("violent", "state_certify", "concede", "criticize_election", "attend_march", "court", "recount", "new_election", "undermine")

for (s in unique(electoral_contestation$survey)) {
  for (v in vars) {
    temp_data <- electoral_contestation %>%
      filter(survey == s) %>%
      select(!!sym(v)) # Use sym() to select the column

    if (nrow(temp_data) > 0) { # Check if there is data to plot
      p <- ggplot(temp_data, aes(x = !!sym(v))) + # Use sym() to map the column
        geom_histogram(aes(y = ..density..), binwidth = 1, fill = "lightblue", color = "black", alpha = 0.5) +
        geom_density(color = "red", bw = 0.5) +
        labs(title = paste("Variable", v, "Survey", s),
             x = v,
             y = "Density") +
        theme_minimal()
      print(p)
    } else {
      print(paste("No data to plot for", v, "in survey", s))
    }
  }
}

```


